### 2025-06-01 conteudo/introduction/intro.tex
add some changes to not be to broad

Our studies and observations reveal that the code duplication problem described
in the previous section cannot be fully addressed by existing tools in
either the formal or grey literature. To fill this gap, we propose an approach
to identify and mitigate code duplications specifically within the Linux
kernel. Our approach includes the development of a tool capable of detecting
duplicated functions and establishing strategies for
addressing these duplications. We tested the approach in the context of two
kernel components, the AMD Display Driver the the Industrial IO subsystem.

@@

In \textbf{Phase I}, we propose, develop, and validate a tool capable of
identifying code duplications at the function level within C codebases. We
expect this tool to effectively detect code duplications in the Linux kernel,
as preliminary investigations suggest that kernel code files often contain a
mix of duplicated and unique functions. To validate the tool, we will use a
triangulation method, comparing its results with established code duplication
database referenced in the literature, and conducting an empirical analysis on
a randomly selected set of files from the AMD Display driver, which our tool
flags as containing duplications. A detailed explanation of our validation
process is provided in Chapter \ref{cha:method}.

Upon completing the first phase, we anticipate having a validated tool for
identifying code duplications within the Linux kernel context. We will then
proceed to the second phase, which focuses on mitigating these
duplications. In \textbf{Phase II}, our objective is to verify if we can 
practically mitigate the duplication finds by the developed tool. 
To achieve this, we have designed an ethnographic approach that enables us to
interact with the Linux development community and gather artifacts that either
support or challenge our proposed mitigation strategies. Figure
\ref{fig:reDesign} summarizes the research steps proposed in this study.

>>>>

Our studies and observations reveal that the code duplication problem described
in the previous section cannot be fully addressed by existing tools in
either the formal or grey literature. To fill this gap, we propose an approach
to identify and mitigate code duplications specifically within the Linux
kernel subsystems, focuding our study on the AMD Display Driver and the Industril IO subsytem. 
Our approach includes the development of a tool capable of detecting
duplicated functions and establishing strategies for
addressing these duplications. We tested the approach in the context of two
kernel components, the AMD Display Driver the the Industrial IO subsystem.

@@

In \textbf{Phase I}, we propose, develop, and validate a tool capable of
identifying code duplications at the function level within C codebases. We
expect this tool to effectively detect code duplications in the targeted subsytems of the Linux kernel,
as preliminary investigations suggest that kernel code files often contain a
mix of duplicated and unique functions. To validate the tool, we will use a
triangulation method, comparing its results with established code duplication
database referenced in the literature, and conducting an empirical analysis on
a randomly selected set of files from the AMD Display driver, which our tool
flags as containing duplications. A detailed explanation of our validation
process is provided in Chapter \ref{cha:method}.

Upon completing the first phase, we anticipate having a validated tool for
identifying code duplications within the context of the Linux kernel subsytems studied. 
We then
proceed to the second phase, which focuses on mitigating these
duplications. In \textbf{Phase II}, our objective is to verify if we can 
practically mitigate the duplication finds by the developed tool. 
To achieve this, we have designed an ethnographic approach that enables us to
interact with the Linux development community and gather artifacts that either
support or challenge our proposed mitigation strategies. Figure
\ref{fig:reDesign} summarizes the research steps proposed in this study.

### 2025-06-01 conteudo/methods/methods.tex
add some changes to not be to broad

This section outlines our approach for mitigating code duplication in the Linux kernel
and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
we present the participant observation research method explored by the author. Then, 
we present the methodolody of the experiment done with the students.

>>>>

This section outlines our approach for mitigating code duplication in the Linux kernel
and describes the ethnographic methodology for engaging with the Linux community. Firstly, 
we present the participant observation research method explored by the author. Then, 
we present the methodolody of the experiment done with the students.

### 2025-06-01 conteudo/results/eval.tex
Add work on 17/05
add some changes to not be to broad

@@

The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
there is a notable increase in function pairs that are not duplications within 
the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
no code duplications were found.

>>>>

\section{Tool Results on BigCloneBench}

@@

The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
there is a notable increase in function pairs that are not duplications within 
the within the context of the AMD Display driver (the Linux kernel component studied here). 
For similarity ranges of 79\%-81\% or higher, there is a high 
percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
no code duplications were found.

### 2025-06-01 conteudo/resumoabstract.tex
gray -> grey
add some changes to not be to broad

O kernel Linux é um projeto de software livre amplamente utilizado, alimentando 96,3\% 
dos um milhão maiores servidores web e 85\% dos smartphones em todo o mundo. Dada a sua extensa 
base de usuários, qualquer adição de recursos, correção de bugs ou tratamento de 
vulnerabilidades de segurança pode impactar milhões de pessoas. Manter o kernel é 
uma tarefa enorme, envolvendo mais de 19 milhões de linhas de código e contribuições 
de mais de mil desenvolvedores. Devido à escala do projeto, nem todas as contribuições 
seguem as melhores práticas, resultando em artefatos de código de baixa qualidade que 
podem complicar a manutenção e o desenvolvimento de novos recursos. Um desses artefatos 
é a duplicação de código, que é o foco desta pesquisa. As ferramentas existentes 
para detectar duplicação de código normalmente comparam dois artefatos de código para 
determinar se são duplicatas. No entanto, essas ferramentas não são adequadas para 
identificar duplicações em bases de código em grande escala, como o kernel Linux, nem 
oferecem orientações sobre como resolver as duplicações detectadas. Apesar das 
pesquisas tanto na literatura formal quanto na cinza, as soluções existentes não 
conseguem resolver os desafios específicos do kernel Linux. Para abordar esse tema, 
esta pesquisa foca em dois objetivos principais. Primeiro, propõe uma nova ferramenta 
para identificar duplicações de código dentro do kernel Linux. Segundo, analisa as 
duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
orientações para futuros contribuidores.

@@

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
96.3\% of the top one million web servers 
and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
is an enormous task, involving more than 19 million lines of code and contributions from over 
thirteen thousand developers. Due to the scale of the project, not all contributions adhere to best 
practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
development. One such artifact is code duplication, which is the focus of this research. 
Existing tools for detecting code duplication typically compare two code artifacts to determine 
if they are duplicates. However, these tools are not well-suited for identifying duplications 
in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
duplications. Despite searches in both formal and gray literature, existing solutions fall 
short of addressing the specific challenges of the Linux kernel. To address this topic, this 
research focuses on two key goals. First, it proposes a new tool to identify code duplications 
within the Linux kernel. Second, it analyzes the identified duplications to extract patterns 
and refactoring methods, offering guidance to future contributors.

>>>>

O kernel Linux é um projeto de software livre amplamente utilizado, alimentando 96,3\% 
dos um milhão maiores servidores web e 85\% dos smartphones em todo o mundo. Dada a sua extensa 
base de usuários, qualquer adição de recursos, correção de bugs ou tratamento de 
vulnerabilidades de segurança pode impactar milhões de pessoas. Manter o kernel é 
uma tarefa enorme, envolvendo mais de 19 milhões de linhas de código e contribuições 
de mais de mil desenvolvedores. Devido à escala do projeto, nem todas as contribuições 
seguem as melhores práticas, resultando em artefatos de código de baixa qualidade que 
podem complicar a manutenção e o desenvolvimento de novos recursos. Um desses artefatos 
é a duplicação de código, que é o foco desta pesquisa. As ferramentas existentes 
para detectar duplicação de código normalmente comparam dois artefatos de código para 
determinar se são duplicatas. No entanto, essas ferramentas não são adequadas para 
identificar duplicações em bases de código em grande escala, como o kernel Linux, nem 
oferecem orientações sobre como resolver as duplicações detectadas. Apesar das 
pesquisas tanto na literatura formal quanto na cinza, as soluções existentes não 
conseguem resolver os desafios específicos do kernel Linux. Para abordar esse tema, 
esta pesquisa foca em dois objetivos principais. 
Primeiro, propõe uma nova ferramenta projetada para identificar duplicações de código dentro 
do kernel Linux, com sua aplicação prática e eficácia avaliadas em subsistemas chave: o AMD 
Display Driver e o subsistema Industrial I/O.
Segundo, analisa as 
duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
orientações para futuros contribuidores.

@@

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
96.3\% of the top one million web servers 
and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
is an enormous task, involving more than 19 million lines of code and contributions from over 
thirteen thousand developers. Due to the scale of the project, not all contributions adhere to best 
practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
development. One such artifact is code duplication, which is the focus of this research. 
Existing tools for detecting code duplication typically compare two code artifacts to determine 
if they are duplicates. However, these tools are not well-suited for identifying duplications 
in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
duplications. Despite searches in both formal and grey literature, existing solutions fall 
short of addressing the specific challenges of the Linux kernel. To address this topic, this 
research focuses on two key goals. 
First, it proposes a new tool designed for identifying code duplications within the Linux kernel, 
with its practical application and effectiveness evaluated in key subsystems: the AMD Display Driver 
and the Industrial I/O subsystem.
Second, it analyzes the identified duplications to extract patterns 
and refactoring methods, offering guidance to future contributors.
